---
title: "FinalExam-2802520665-Davina Oktaviani"
author: "Davina Oktaviani"
date: "2025-06-24"
output:
  html_document:
    toc: true
    toc_depth: 2
    code_folding: hide
    df_print: paged
    fig_width: 5
    fig_height: 7
    dev: png
    theme: readable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```


#Video Explanation
Youtube: you can watch the video [here](https://youtu.be/vtJB7W2gdl8?si=zbrop3Vm3fwwnPl7)
Google Drive: you can also watch the video [here](https://drive.google.com/file/d/1KORu7753zwBs9yCUrpGkorsRH0fUItZy/view?usp=drivesdk)



```{r}
library(tidyverse)
library(plotly)
library(dplyr)
library(vcd)
library(tidyr)
```

#Read data
```{r}
df = read.csv("data_B.csv")
head(df)
```



## 1. [LO1, LO2 – 30 Points] Data Quality Assessment and Preprocessing {.tabset}

### a. [LO1 - 10 Points] Identify data quality issues 

#### Missing values

```{r}
df[df == "unknown"] = NA
df[df == "NA"] = NA
```

```{r}
numeric_cols = c("age", "duration", "campaign", "pdays", "previous")
df[numeric_cols] = lapply(df[numeric_cols], function(x) as.numeric(as.character(x)))
```


```{r}
df = df %>% drop_na(-pdays)
summary(df)
```

#### Duplicate records
```{r}
df = df[!duplicated(df), ]
nrow(df)
```

#### Inconsistent or unknown categories 
```{r}
factor_cols = c("job", "education", "marital", "default", 
                 "housing", "loan", "contact", 
                 "month", "day_of_week", "poutcome", "y")
df[factor_cols] = lapply(df[factor_cols], as.factor)
str(df[factor_cols])
```


#### Outliers or extreme values
```{r}
handle_outliers = function(df, col) {
  Q1 = quantile(df[[col]], 0.25, na.rm = TRUE)
  Q3 = quantile(df[[col]], 0.75, na.rm = TRUE)
  IQR = Q3 - Q1
  lower = Q1 - 1.5 * IQR
  upper = Q3 + 1.5 * IQR
  df = df[df[[col]] >= lower & df[[col]] <= upper, ]
  return(df)
}

df = handle_outliers(df, "age")
df = handle_outliers(df, "duration")
df = handle_outliers(df, "campaign")
df = handle_outliers(df, "previous")

summary(df[c("age", "duration", "campaign", "previous")])
```


#### Noisy or irrelevant values
```{r}
df$pdays_cat = ifelse(df$pdays == 999, "never_contacted", "contacted_before")
df$pdays_cat = as.factor(df$pdays_cat)
table(df$pdays_cat)
```


### b. [LO1, LO2 - 10 Points] Handle the issues you identified

#### Clean the data based on your earlier findings.
The data has been cleaned based on the issues identified earlier, including missing values, duplicate records, inconsistent categories, outliers, and noisy values

#### Explain your cleaning decisions (e.g., whether you imputed, removed, or recoded values). 
- Missing values: Replaced "unknown" and "NA" strings with NA, then removed rows with missing values, excluding pdays to retain meaningful data.
- Data types: Converted numeric columns from character to numeric for accurate analysis.
- Duplicates: Removed duplicate rows to avoid bias.
- Categories: Converted categorical columns to factors for consistent handling.
- Outliers: Removed outliers using the IQR method on numeric columns, except pdays, since 999 is meaningful.
- Noisy values: Kept pdays == 999 (not treated as missing), and created pdays_cat to label clients as "never_contacted" or "contacted_before".
- Output: Cleaned dataset saved as clean_data.csv.


#### Create a cleaned version of the dataset called clean_data. 
```{r}
clean_data = df
write.csv(clean_data, "clean_data.csv", row.names = FALSE)
head(clean_data)
```


### c. [LO1, LO2 -10 Points] Summary statistics and initial exploration 

#### Calculate central tendency and variability for numerical columns
```{r}
summary(clean_data %>% select_if(is.numeric))
```

#### Create a summary table to explore one variable across categories (e.g., age by job type)

##### age by job
```{r}
clean_data %>%
  group_by(job) %>%
  summarise(mean_age = mean(age, na.rm = TRUE),
            median_age = median(age, na.rm = TRUE),
            n = n()) %>%
  arrange(desc(n))
```


##### duration by education
```{r}
clean_data %>%
  group_by(education) %>%
  summarise(
    mean_duration = mean(duration, na.rm = TRUE),
    median_duration = median(duration, na.rm = TRUE),
    n = n()) %>%
  arrange(desc(mean_duration))
```

##### campaign by marital
```{r}
clean_data %>%
  group_by(marital) %>%
  summarise(
    mean_campaign = mean(campaign, na.rm = TRUE),
    median_campaign = median(campaign, na.rm = TRUE),
    n = n())
```

#####  pdays by contact
```{r}
clean_data %>%
  group_by(contact) %>%
  summarise(
    mean_pdays = mean(pdays, na.rm = TRUE),
    median_pdays = median(pdays, na.rm = TRUE),
    n = n())
```

##### duration by month
```{r}
clean_data %>%
  group_by(month) %>%
  summarise(
    mean_duration = mean(duration, na.rm = TRUE),
    median_duration = median(duration, na.rm = TRUE),
    n = n())
```

##### age by education
```{r}
clean_data %>%
  group_by(education) %>%
  summarise(
    mean_age = mean(age, na.rm = TRUE),
    median_age = median(age, na.rm = TRUE),
    n = n())
```



## 2. [LO1, LO2, LO3 – 35 Points] Relationship Analysis {.tabset}

### a. [LO1 - 10 Points] Explore variable relationships. Choose three variables that might influence the campaign outcome (y)

I selected 3 variables that might influence the campaign outcome(y):
1. age
2. job
3. duration

#### Use correlation (for numeric), cross-tabulation (for categorical), or grouped summaries.

```{r}
clean_data$y_binary = ifelse(clean_data$y== "yes", 1, 0)
```

##### age vs y_binary (both are numeric)
```{r}
shapiro.test(sample(clean_data$age))
shapiro.test(sample(clean_data$y_binary))
```
The result shows that the p-value for both variables is less than 0.05, indicating that the variables do not follow a normal distribution.

```{r}
cor.test(clean_data$age, clean_data$y_binary, method="kendall")
```
Since the p-value is also below 0.05, there is a statistically significant relationship between age and y_binary.


##### job vs y (both are categorical)
```{r}
t1=table(clean_data$job, clean_data$y)
t1
```
since the contingency table contains cells with values less than 5, we use Cramer's V test.

```{r}
library(vcd)
assocstats(t1)
```
based on the likelihood ratio test, the p-value is less than 0.05, indicating a significant association between job and y.


##### duration vs y_binary (numeric and categorical)
```{r}
model = lm(y_binary~duration, data = clean_data)
summary(model)
```
the p-value is less than 0.05, meaning that the duration variable has a significant effect on y_binary.

#### Discuss any patterns or trends you observe.
- The Kendall correlation test between age and y_binary produced a p-value < 0.05, indicating a statistically significant relationship.
- The association between job and y was found to be significant (p < 0.05).
- The linear model between duration and y_binary showed a significant p-value, indicating that call duration affects the likelihood of subscription.


### b. [LO2, LO3 -10 Points] Feature engineering. Create 3 new variables to enhance your analysis:

#### age_group with levels: "<30", "30–50", ">50"
```{r}
clean_data = clean_data %>%
  mutate(age_group = case_when(
    age < 30 ~ "<30",
    age >= 30 & age <= 50 ~ "30–50",
    age > 50 ~ ">50"
  ))

table(clean_data$age_group)
```

#### recency_level based on pdays: "never", "recent", "old" 
```{r}
clean_data = clean_data %>%
  mutate(recency_level = case_when(
    pdays == 999 ~ "never",
    pdays <= 5 ~ "recent",
    pdays > 5 & pdays < 999 ~ "old",
  ))

table(clean_data$recency_level)
```


#### total_contact_score = campaign + previous
```{r}
clean_data = clean_data %>%
  mutate(total_contact_score = campaign + previous)

summary(clean_data$total_contact_score)
```


### c. [LO2, LO3 - 15 Points] Interactive visualizations of relationships. Create three interactive charts using Plotly: 

#### A chart comparing two categorical variables 
```{r}
plot_ly(clean_data %>% count(job, y),
        x = ~job, y = ~n, color = ~y, type = 'bar', text = ~n, textposition = 'auto') %>%
  layout(title = "Campaign Outcome by Job",
         xaxis = list(title = "Job"),
         yaxis = list(title = "Number of Clients"),
         barmode = "stack")
```


#### A chart comparing one numeric and one categorical  
```{r}
plot_ly(clean_data, x = ~y, y = ~age, type = 'box', color = ~y) %>%
  layout(title = "Age Distribution by Campaign Outcome",
         xaxis = list(title = "Campaign Outcome"),
         yaxis = list(title = "Age"))
```


#### A multi-dimensional chart, e.g., faceted by month or grouped by education
```{r}
plot_ly(clean_data, x = ~education, y = ~total_contact_score, color = ~y, type = 'box') %>%
  layout(title = "Total Contact Score by Education and Outcome",
         xaxis = list(title = "Education"),
         yaxis = list(title = "Total Contact Score"))
```


## 3. [LO1, LO3, LO4 – 25 Points] Insight Storytelling and Recommendations. You are now presenting your insights to a team of non-technical stakeholders. {.tabset}

### a. [LO3 - 10 Points] Data storytelling through visual design. Choose 2 key findings and design an interactive chart with excellent visual storytelling. principles: Clean title, Informative labels, Highlight key differences using color or annotation

#### age_group vs y
```{r}
plot_ly(clean_data %>% count(age_group, y),
        x = ~age_group, y = ~n, color = ~y, type = 'bar', text = ~n, textposition = 'auto') %>%
  layout(title = "Response Rate by Age Group",
         xaxis = list(title = "Age Group"),
         yaxis = list(title = "Number of Clients"),
         barmode = "stack")
```


#### poutcome vs y
```{r}
plot_ly(clean_data %>% count(poutcome, y),
        x = ~poutcome, y = ~n, color = ~y, type = 'bar', text = ~n, textposition = 'auto') %>%
  layout(title = "Campaign Success Based on Previous Outcome",
         xaxis = list(title = "Previous Outcome"),
         yaxis = list(title = "Number of Clients"),
         barmode = "stack")
```


### b. [LO1 - 10 Points] Interpret your findings. In 3–5 sentences, write a clear narrative explaining: Central trends (e.g., median age, contact duration), Any unusual values or segments (e.g., age group with lowest response), Possible correlations or causality risks

- The average client age is 39 years, with a median of 37, mostly concentrated between ages 35–40.
- Students are the youngest group, while retirees are the oldest. Other job types are mostly aged 35–45.
- Clients with "illiterate" education have the highest average call duration, while those in professional courses have the lowest.
- Married clients are contacted more frequently than single or divorced ones.
- The highest average contact duration occurs in December.
- The "admin" job category has the highest proportion of positive (yes) responses.
- The median age of clients who responded "yes" is lower than those who responded "no", indicating that younger clients tend to be more responsive.
- Clients with "basic.6y" education show the highest number of positive responses.
- The age group 30–50 shows the highest proportion of "yes" responses.
- High contact duration may be caused by the client's initial interest, suggesting the possibility of reverse causality.


### c. [LO1, LO4 - 5 Points] Make business recommendations. Write two specific, actionable business recommendations based on your findings. 

- Focus the campaign on clients aged 30–50 and those with admin jobs, as they show the highest "yes" response rates.
- Prioritize clients with a history of successful previous campaigns (poutcome = success), as they are more likely to respond again.
- Ensure that call durations last at least 2 minutes, as clients who respond "yes" generally have longer contact durations than those who do not.